### 2010-02-24 bgps_pipeline/methods/ms.tex

% Figure list
% 1) fig:Coverage
% 2) fig:BGPSMontage
% 3) fig:Bandpass
% 4) fig:Array
% 5) fig:CalibrationCurves
% 6) fig:Flatfield
% 7) fig:CygnusComparison
% 8) fig:PointingCalibrators
% 9) fig:PointingModel
% 10) fig:SCUBAPointingComparison
% 11) fig:IterativeMapping
% 12) fig:Flagging
% 13) fig:GlitchDistribution
% 14) fig:PCA_Graphical
% 15) fig:Deconvolution
% 16) fig:Noise
% 17) fig:NoiseVsLongitude
% 18) fig:PCA_Filter
% 19) fig:AperturePhotometry
% 19) fig:sampleimages

% Major Sections
% sec:Introduction
% sec:Observations
% sec:FluxCalibration
% sec:Astrometry
% sec:Mapping
% sec:Noise
% sec:Photometry
% sec:FinalMaps
% sec:Discussion

% calibration
% you should look @ calibration/plot_dcfluxes.pro in my publish/ directory, and /usb/scratch1/planets/planet_dcfluxes.txt is the actual raw data
%/home/milkyway/student/ginsbura/bgps_pipeline/calibration/plot_dcfluxes.pro?

% glitches
% I can't send you raw data for the glitches, but the best way I can think of dealing with them is taking one of my postiter.sav files in some directory and grabbing ac_bolos[where(flags)]...... except that include hand-flagged stuff too

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                       % 
%       Bolocam Galactic Plane Survey Description       %
%                                                       % 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[12pt,preprint]{aastex}
\documentclass{emulateapj}

%%% THIS ACHIEVES THE HOLY GRAIL OF 1 IN MARGINS ALL AROUND!!! %%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\oddsidemargin  0.0in
	\evensidemargin 0.0in
	\textwidth      6.5in
	\headheight     0.0in
	\topmargin      0.375in
	\textheight=9in
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{rotating}
\usepackage{subfigure}
\usepackage{datetime}
\usepackage{fmtcount}

\newtimeformat{edttime}{\twodigit{\THEHOUR}:\twodigit{\THEMINUTE}~EDT}
\settimeformat{edttime}
\newdateformat{eurodate}{\dayofweekname{\THEDAY}{\THEMONTH}{\THEYEAR} \THEDAY \monthname[\THEMONTH] \THEYEAR}

\citestyle{aa}

\bibliographystyle{apj_w_etal}

\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\xad}{\vect{x}}

\newcommand{\ammonia}{\mbox{{\rm NH}$_3$}}
\newcommand{\form}{H$_2$CO}
\newcommand{\td}{\mbox{$T_D$}}
\newcommand{\jj}[2]{\mbox{$J = #1\rightarrow#2$}}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\def\Check#1{{\tt #1}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}
\newcommand{\lon}{\ensuremath{l}}
\newcommand\jyb{Jy~beam$^{-1}$}

\newcommand{\epsi}{\varepsilon}

% -----------------------------------------------------------------------------
% BGPS specific definitions
\def\bgps{BGPS}
\newcommand{\bgpsarea}{150}
\newcommand{\bcamfwhm}{33\arcsec}
\newcommand{\ncores}{7628}
\newcommand{\bgpsdepthlow}{20}
\newcommand{\bgpsdepthhigh}{50}
% -----------------------------------------------------------------------------
% Missions
\def\spitzer{{\em Spitzer}}
% -----------------------------------------------------------------------------

\newcommand{\TBD}{{\bf TBD}}

\def\Figure#1#2#3#4{
\begin{figure}[htb]
\epsscale{#4}
\plotone{#1}
\caption{#2}
\label{#3}
\end{figure}
}

\def\FigureTwo#1#2#3#4#5{
\begin{figure}[htb]
\epsscale{#5}
\plottwo{#1}{#2}
\caption{#3}
\label{#4}
\end{figure}
}

\def\Table#1#2#3#4#5{
\begin{deluxetable}{#1}
\tablewidth{0pt}
\tablecaption{#2}
\tablehead{#3}
\startdata
\label{#4}
#5
\enddata
\end{deluxetable}
}

\newcommand{\penn}{1}
\newcommand{\casa}{2}
\newcommand{\utexas}{3}
\newcommand{\virginia}{4}
\newcommand{\ucf}{5}
\newcommand{\wisc}{6}
\newcommand{\jpl}{7}
\newcommand{\ubc}{8}
\newcommand{\hawaiihilo}{9}
\newcommand{\hawaiimain}{10}
\newcommand{\arizona}{11}

	% End definitions

\slugcomment{DRAFT: \currenttime\ \eurodate\today}

\shorttitle{BGPS}
\shortauthors{Aguirre et al.}

\begin{document}

\newcounter{subfig}
\setcounter{subfig}{1}

\title{The Bolocam Galactic Plane Survey: Survey Description and Data
Reduction}

%\newpage
%\pagebreak
%
%\tableofcontents
%
%\newpage
%\pagebreak
%

\author{James Aguirre\altaffilmark{\penn},
        Adam Ginsburg\altaffilmark{\casa},
        Miranda K. Nordhaus\altaffilmark{\utexas},
	Meredith Drosback\altaffilmark{\virginia},
        John Bally\altaffilmark{\casa}, 
	Cara Battersby\altaffilmark{\casa},
	Eric Todd Bradley\altaffilmark{\ucf},
%        Richard Chamberlin\altaffilmark{\cso},
	Claudia Cyganowski\altaffilmark{\wisc},
	Darren Dowell\altaffilmark{\jpl},
	Neal J. Evans II\altaffilmark{\utexas},
        Jason Glenn\altaffilmark{\casa},
        Paul Harvey\altaffilmark{\utexas,\casa},
        Erik Rosolowsky\altaffilmark{\ubc},
%        Wayne Schlingman\altaffilmark{9},
%        Yancy Shirley\altaffilmark{9},
        Guy S. Stringfellow\altaffilmark{\casa},
        Josh Walawender\altaffilmark{\hawaiihilo}, and 
        Jonathan Williams\altaffilmark{\hawaiimain}
}

\affil{{$^\penn$}{\it{Department of Physics and Astronomy, University of
      Pennsylvania, Philadelphia, PA }}}
\email{\myemail}

\affil{{$^\casa$}{\it{CASA, University of Colorado, CB 389-UCB, Boulder, CO 80309}}}
%, \email{John.Bally@casa.colorado.edu}
%      \email{adam.ginsburg@colorado.edu}
%      \email{cara.battersby@colorado.edu}
%      \email{Jason.Glenn@colorado.edu} \email{pmh@astro.as.utexas.edu}
%      \email{Guy.Stringfellow@colorado.edu} }}

\affil{{$^\utexas$}{\it{Department of Astronomy, University of Texas,
      1 University Station C1400, Austin, TX 78712}}}
%,            
%      \email{nje@astro.utexas.edu},
%      \email{nordhaus@astro.utexas.edu}}}

\affil{{$^\virginia$}{\it{Department of Astronomy, University of
Virginia, P.O. Box 400325, Charlottesville, VA 22904}}}
%, 
%    \email{meredith.drosback@origins.colorado.edu}}}

\affil{{$^\ucf$}{\it{Department of Physics, University of Central
Florida}}}
%,           
%                                              } ,   \email{tbradley@physics.ucf.edu}} }

\affil{{$^\wisc$}{\it{Department of Astronomy, University of Wisconsin,
       Madison, WI 53706}}}
% , \email{cyganow@astro.wisc.edu}}}

\affil{{$^\jpl$}{\it{Jet Propulsion Laboratory, California Institute
   of Technology, 4800 Oak Grove Dr., Pasadena, CA 91104}}}
%,
%       \email{cdd@submm.caltech.edu}}}

\affil{{$^{\ubc}$}{\it{Department of Physics and Astronomy, University
       of British Columbia, Okanagan }}}
% ,               \email{erik.rosolowsky@ubc.ca}}}

%\affil{{$^{\arizona}$}{\it{ Steward Observatory, University of
%Arizona, 933 North Cherry Avenue, Tucson, AZ 85721 }}}
%%\email{yshirley@as.arizona.edu} \email{wschlingman@as.arizona.edu}}}

\affil{{$^{\hawaiihilo}$}{\it{Institute for Astronomy, University of
        Hawaii,640 N. Aohoku Pl., Hilo, HI 96720}}}
%\email{joshw@ifa.hawaii.edu} }}

\affil{{$^{\hawaiimain}$}{\it{Institute for Astronomy, University of
Hawaii,680 Woodlawn Drive, Honolulu, HI 96822}}}
%\email{jpw@ifa.hawaii.edu}}

%\altaffiltext{\penn}{University of Pennsylvania, Philadelphia, PA 19104}
%%Jansky Fellow, National Radio Astronomy Observatory} 
%\altaffiltext{\casa}{CASA, University of Colorado, CB 389, Boulder, CO 80309}
%\altaffiltext{\utexas}{Department of Astronomy, University of Texas, 
%       1 University Station C1400, Austin, TX 78712}
%\altaffiltext{4}{Caltech Submillimeter Observatory, Hilo, HI}
%\altaffiltext{5}{Institute for Astronomy (IfA), University of Hawaii
%640 N. Aohoku Pl., Hilo, HI 96720}
%
\begin{abstract}

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm
continuum survey of \bgpsarea\ square degrees of the Galactic Plane
with 130 in the 1st quadrant, 31 in the 2nd, and 6 in the 3rd.  The
BGPS represents the first large area, systematic survey of the
Northern Galactic Plane in the millimeter continuum without
pre-selected targets.  The survey has detected \ncores\ clumps to a
limiting non-uniform 1-$\sigma$ noise level in the range
\bgpsdepthlow\ and \bgpsdepthhigh\ m\jyb RMS.  We present maps of 97
contiguous square degrees from $\lon=-10.5$ to $\lon=86.5$ and
$b=\pm0.5$ with a resolution of $\sim$\bcamfwhm, as well as regions in
IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most striking
feature of the maps is the filamentary structure of millimeter
emission throughout the Galaxy.

This paper details the survey observations and data reduction methods.
We treat carefully the determination of pointing and flux density
calibration errors and compare our results to the literature.  Data
processing algorithms that separate astronomical signals from
time-variable atmospheric flux density in the data time-stream are
discussed.  The algorithms reproduce the structure of the astronomical
sky over a limited range of angular scales and produce artifacts in
the vicinity of bright sources. Specifically, extended emission smooth
on scales larger than about 3\arcmin\ is not fully recovered in the
images.  The spatial frequency response function of the Bolocam
Galactic Plane Survey is discussed.

This paper serves as a companion and guide to the public data release
through NASA's Infrared Processing and Analysis Center (IPAC).  These
images represent an important finder chart for future observations
with ALMA and a long-wavelength spectral band complementary to
ATLASGAL and imminent submillimeter surveys with SCUBA-2 and
Herschel-SPIRE.

\end{abstract}

\keywords{
ISM: - molecular clouds --
stars: formation -- high mass
millimeter continuum
}

%\include{bgps_methods_introduction}

\section{Introduction}
\label{sec:Introduction}

Millimeter-wavelength continuum surveys of the Galactic plane provide
the most efficient way to find molecular clumps that are the likely
sites of formation for massive stars and star clusters. The
development of detector arrays has made blind surveys of large areas
possible.  Such surveys bypass the need for selection based on the
presence of embedded stars or star clusters, infrared sources, masers,
or radio continuum emission. In particular, they can locate molecular
clumps before stars form, providing vital information on the initial
conditions of star formation.  These surveys can also provide valuable
constraints on the physical properties of the clumps, especially
masses and mean densities, in conjunction with distance information.

Galaxy-wide surveys are essential for measuring the impacts of the
environment on clump properties and star formation activity.  Do clump
properties vary with Galactocentric distance, or with location with
respect to spiral arms?  Do they depend on the level of nearby star
formation activity?  Answering these questions in our Galaxy will
provide the essential ``ground truth'' required for the analysis of
distant galaxies where individual clouds and clumps are not resolved
and only galaxy-wide average quantities can be measured.

Surveys of the galaxy in CO line emission \citep{dame01,jackson06}
have identified the locations of large molecular clouds and traced the
transition from atomic gas in the outer galaxy to molecular gas in the
inner galaxy.  Together with kinematic distances
\citep[e.g.][]{RotationCurve} and, more recently, accurate distances
from VLBI \citep{reid09}, these surveys have begun to reveal the
spiral structure of the Milky Way. As this picture comes into sharper
focus, we can begin to address the old question of the relationship
between spiral density waves and star formation.

To pursue this question, we need tracers of higher density material.
Studies of nearby clouds have demonstrated that star formation is far
from uniform over molecular clouds, but is concentrated in unusually
dense regions called clumps or cores, which fill a small fraction of
the area of most clouds \citep{lada91,enoch07,evans09}.
%[added ref to Lada and changed ref to Enoch to the summary paper. NJE]
Following \citet{williams00} and \citet{mckee07}, we use
the term `core' to refer to a very dense region destined to form an individual
star or small multiple star system and the term `clump' to refer to a region
likely to form a group or cluster of stars. While clumps forming massive
clusters may be even denser than cores forming low mass stars 
\citep{mueller02}, in general, clumps are larger and have a lower mean 
density \citep{mckee07}. Clumps can be
traced with line emission from molecules such as \ammonia, \form, CS, HCN, etc.
Spectral lines also provide excellent diagnostics of line-of-sight motions,
temperatures, and densities.  However, the interpretation of these gas tracers
is complicated by variations in tracer abundances caused by freeze-out onto
grains, sublimation caused by star formation, and subsequent
complex chemical processing \citep{vandishoeck98}. Interpretation is
further complicated by uncertainties in optical depths and excitation
conditions and the impacts of radiation fields and shocks. These complications
make the derivation of column densities, masses, and other physical properties
of star-forming clumps very difficult.  A limitation of spectral line surveys
is that the tracers of denser gas tend to be weak and multi-element receiver
arrays have only a modest number of elements, making blind surveys very
time-consuming. Consequently, no blind surveys of the Galaxy have been done
in lines that trace the molecular clumps.

In contrast, focal-plane arrays containing hundreds of individual
bolometers sensitive to millimeter and sub-millimeter (sub-mm)
radiation are now available. This advance in technology enables blind
surveys of the Galactic plane, providing a uniform inventory of
massive star-forming and starless clumps.  Catalogs resulting from
these surveys will provide the data base for subsequent observations
in tracers such as hard-to-excite molecular lines and higher
resolution studies with instruments like CARMA and ALMA.  A
significant advantage of continuum observations of dust at long
wavelengths is the low optical depth of the dust \citep{johnstone06},
permitting a straight-forward estimate of the mass of the emitting
region.

Submillimeter observations of Galactic sources have been conducted by
a variety of different groups focusing on small regions.  The
Submillimeter Common-User Bolometer Array (SCUBA) on the 15-m James
Clerk Maxwell Telescope mapped many significant regions in the
galactic plane over its 8 year lifetime
\citep{holland99,difrancesco08}, but never completed a contiguous
survey of the galactic plane.  The Bolocam instrument has been used on
the 10.4-m Caltech Submiliimeter Observatory (CSO) to map nearby large
molecular clouds at 1.1 mm and identify hundreds of cores
\citep{enoch06}.  MAMBO on the 30-m IRAM telescope has been used at
1.2 mm to map both small, nearby cores \citep{kauffmann08} and more
distant regions of more massive star formation \citep{motte07}.  None
of these surveys covered a significant fraction of the Galactic plane.

Two large surveys of the Galactic plane are in progress.  The ATLASGAL
survey, using the LABOCA instrument at 870 \micron\ with 12\arcsec\
resolution on the APEX telescope in Chile, will cover the southern
Galactic plane and some of the north.  Initial results from this
survey have been reported by \citet{schuller09}. The overlap region
between BGPS and ATLASGAL is in the region $-10.5 < \lon < 11.5$ and
$15 < \lon < 21$, $-0.5 < b < 0.5$, though ATLASGAL will eventually
expand up to $\lon = 60$.  The two surveys provide complementary
samples of the long-wavelength end of the spectral energy distribution
in Galactic star forming regions.

This paper presents the results of the Bolocam Galactic Plane Survey
(BGPS).  The survey covers the contiguous range $-10.5 < \lon < 86.5,
-0.5 < b < 0.5$ with additional coverage to higher $|b|$ in the range
$77 < \lon < 86.5$ and in the $\lon=111$, W3/4/5, Gem OB1, and IC 1396
regions.  It is complementary to the survey by \citet{schuller09} in
that it targets the part of the plane visible from the northern
hemisphere and uses a somewhat longer wavelength, allowing comparison
of properties in the region of overlap with the ATLASGAL survey.  In
addition, the BGPS covers the Galactic anti-center region, allowing
consistent comparison of very different parts of the Galaxy.  At 870
\micron, $h\nu/k = 16.5$ K, compared to 13.1 K at 1.1 mm.  Because
starless clumps may be quite cold ($\td \approx 10$ K), the
uncertainty in mass caused by uncertain temperature is slightly less
at 1.1 mm.  For example, if the actual $\td = 20$ K, rather than 10 K,
the mass estimate would decrease by a factor of 3.3 at 870 \micron\
versus a factor of 2.9 at 1.1 mm.  In addition, the Bolocam bandpass
excludes the CO \jj{2}{1} line, which can be a significant contaminant
on top of the dust continuum emission.

We have used the 144-element Bolocam focal plane array, mounted at the
Cassegrain focus of the CSO, to survey over \bgpsarea\ square degrees
of the northern Galactic plane at a wavelength of 1.1 mm in the dust
continuum.  Figure \ref{fig:Coverage} shows the coverage of the \bgps.
%[Note: ref to figure 2 with our images is commented, needs to be called
%in Section 9. NJE]
We have detected \ncores\ clumps in the surveyed fields, providing a
sample of clumps suitable for multi-wavelength and high-resolution
studies with existing telescopes and future facilities.  This sample
is free of the usual biases associated with observing only regions
with signposts of star formation, such as HII regions or masers. While
sensitivity limits will bias the survey against distant, low-mass
objects, follow-up studies indicate that many sources are more
massive, distant objects \citep{bally09,nordhaus09, rosolowsky09}.
%[refs to papers in prep by John, Miranda, and Erik, where distance is
%discussed a bit. Erik's last version had more discussion of this
%topic, so need to coordinate with that. NJE] 
These objects are likely to be aggregates of star forming cores found
in a larger molecular cloud environment, which we will refer to as
clumps \citep{williams00}.  The clump catalog is described in a
companion paper \citep{rosolowsky09}.

The outline of the paper is as follows. Section \ref{sec:Observations}
describes the instrument and the observations.  Section
\ref{sec:Overview} briefly describes the data reduction process, while
Sections \ref{sec:FluxCalibration} to \ref{sec:FinalMaps} provide
greater detail on flux calibration, astrometry, the mapping algorithm,
noise, photometry, and the public data release.  We conclude in
Section \ref{sec:Discussion} with brief discussion of the immediate
implications of the emission revealed by the BGPS.

% TEST
% Useful, but not here
%\begin{figure}
%  \subfigure[$\lon=-10.5$ to $\lon=19.5$ Units are \jyb.  The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.  The noise is more pronounced from $\lon=-7$ to $\lon=-2$
%	because this region was observed less.]{\label{fig:BGPSMontage-a}\includegraphics[scale=0.75]{fig1_pg1_bw_inv_bar_label}}\\
%\pagebreak
%\subfigure[$\lon=19.5$ to $\lon=49.5$.  Units are \jyb .  G34.3+0.15, W
%	51, W 43, W 49, and M 17 appear to be saturated, but this is
%	only a display artifact.]
%{\label{fig:BGPSMontage-b}\includegraphics[scale=0.75]{fig1_pg2_bw_inv_bar_label}}
%
%\end{figure}
%% --------------------------------
%

\section{Observations}
\label{sec:Observations}

Bolocam\footnote{{\tt http://www.cso.caltech.edu/bolocam}} is the
facility 144-element bolometer array camera operating on the 10.4 m
dish of the Caltech Submillimeter Observatory (CSO) on the summit of
Mauna Kea.  We used the filter configuration with a bandcenter of
268~GHz (hereafter 1.1~mm) and fractional bandwidth $\Delta \nu/\nu =
0.17$ (45.5 GHz).  The passband is constructed to exclude the
$^{12}\mathrm{CO}(2 \to 1)$ emission line.  Figure \ref{fig:Bandpass}
shows the passband.  We compute color corrections in Appendix
\ref{app:ColorCorrections}.

The array field-of-view is 7\arcmin.5, with individual detectors
having nearly Gaussian beams of \bcamfwhm\ FWHM.  The spacing of the
pixels at 1.1 mm is 1.6$f\lambda$, so the focal plane is not
instantaneously sampled.  The Bolocam instrument is described in
greater detail in \citet{haig04} and \citet{glenn03}.

%\subsection{Observations}

The maps presented here were acquired during six separate observing
sessions at the CSO over the course of two years.  The observing
epochs are given in Table \ref{tab:Observing}, along with
corresponding ranges for the zenith opacity $\tau_{225}$ of the CSO
tipper tau.  Bolocam observations were typically scheduled when
$\tau_{225} > 0.06$.  Between observing epochs, Bolocam was removed
from its mount at the re-imaged Cassegrain focus and stored warm.
Thus the flux calibration and pointing model must be re-computed for
each epoch to allow for variations in the instrument and optics.  We
found that the flux calibration did not in fact differ significantly
between epochs.  We observed additional pointing sources during Epoch
V and therefore used it as the master pointing model.

%\subsection{Observing Strategy}

% xraster 120 11280 23 /step_size=162 /equatorial /position_angle=36.753118 /alternate_direction /rot /settling_time = 03

%[Reword section?] 
Our basic observing strategy was to raster scan Bolocam by moving the
primary mirror of the CSO to modulate the astrophysical signal faster
than fluctuations in atmospheric opacity.  Each field was scanned with
alternating raster scans along $l$ and $b$.  In Epoch I-II, the
fundamental scan unit was a $1\arcdeg \times 1\arcdeg$ block, scanned
23 times in \lon\ and $b$.  Starting with Epoch III, the fundamental
scan unit was changed to a $3\arcdeg \times 1\arcdeg$ block for
increased mapping efficiency.  Each block was scanned with 23 scans
along lines of constant $b$, and 67 scans along lines of constant $l$.
In both cases, the spacing between adjacent rasters was 162\arcsec.
The total time for such a scan was 39 (48) minutes for the $b$ ($l$)
scans.  The data were electronically sampled at 10 Hz along the scan
direction, slightly higher than the Nyquist rate for the scan speed of
120\arcsec\ s$^{-1}$.

In order to fill in the instanteously undersampled focal plane, we
began using in Epoch II a field rotator to mechanically align the
focal plane at an optimal orientation along the scan direction to
improve the sampling orthogonal to the scan direction.  This is shown
in Figure \ref{fig:Array}.  Achieving good sampling is complicated by
a number of missing bolometers.  A simulation was performed to
determine the optimal angle to rotate the array with respect to the
scan direction to account for the beam spacing and missing bolometers.
During the turnaround following a raster, the field rotator is
adjusted to this optimum angle.
%CHECK
% per Adam, 2009/3/26
Without the field rotator, the coverage shows variations of 100\% from
pixel to pixel in a single raster of a field.  With the rotator, this
is reduced to $\sim40\%$.

\section{Overview of the Data Reduction}
\label{sec:Overview}

A number of different descriptions of the reduction of Bolocam data
have been published \citep{laurent05,enoch06,sayers09}.  This work
addresses the issues specific to the BGPS, significantly improving the
method of \citet{enoch06} for generating maps and characterizing their
properties and is similar in many respects to the methods presented in
\citet{cotton09} or \citet{kovacs08}.  It incorporates some of the
technical developments of \citet{sayers09}, though the exquisite
control of systematics and noise of that analysis is neither necessary
nor possible here.

The data were reduced with a custom pipeline written specifically for
the BGPS.  A key element of the reduction is the iterative estimation
of the atmospheric fluctuations and the astrophysical signal.  The
atmospheric model is developed from a set of principal components of
the bolometer signal time series under the assumption that the bulk of
the correlated signal is atmospheric (under this assumption,
astrophysical extended structures are treated as atmospheric noise).
The iterative estimation process then helps to restore correlated flux
due to astrophysical emission (Section \ref{sec:Mapping}).

The Bolocam observations are sensitive to structure between \bcamfwhm\
and about 200\arcsec\ (half the array field of view, see figure
\ref{fig:PCA_Filter}).  We present a detailed analysis of the filter
function in our survey in figure \ref{fig:PCA_Filter}.  The data were
sampled onto a uniform 7\farcs2 grid.  Flux density calibrations were
performed using observations of Mars and Uranus, and the overall
calibration accuracy is estimated to be $\sim$15\%.  The flux
calibration is described in section \ref{sec:FluxCalibration}.

%A calibration scheme capable of
%5\% systematic and $<$5\% statistical errors is being implemented as outlined in
%\citet{Sayers2009}.
%To account for offsets between the observations, each observation was mapped
%individually and aligned to a `master' field that has been tied to
%pointing calibrators. 
Noise properties across the survey are not completely uniform, but the
measured $1\sigma$ level is $\approx43\pm7$ m\jyb within the
$b=\pm.5$ portion of the plane in the 1st quadrant.
%After the pipeline mapping, the fields were combined using the COMBINE
%command in IRAF with WCS offsets.

A separate model was constructed for each epoch, but epoch V was used
as the primary pointing calibration and is shown in figure
\ref{fig:PointingModel}.  In epoch V and VI, pointing sources were
observed more frequently to obtain a better-constrained master
pointing model. The absolute pointing accuracy was $\sim6\arcsec$ RMS.
The Epoch V pointing model was used to make pointing reference fields
to which all other fields were aligned using a cross-correlation
algorithm (Section \ref{sec:Astrometry}).

% A single $3^\circ \times
% 1^\circ$ observation of each region was registered to the pointing calibrators
% as a ``master field'', and additional observations of that field were aligned
% to the master (Section \ref{sec:Astrometry}).

Bright and large sources leave an impact on their surroundings when
cleaned with a principal component subtraction.  \object{Sgr B2},
\object{G 34.3+0.15}, and other sources with fluxes $\geq 10$ Jy are
surrounded by substantial negative bowls.  Some flux is restored in
these bowls using the iterative process, but around the brightest
sources the impact of the source on the background estimation is too
great to be fully recovered (Section \ref{sec:Photometry}).

Instrument glitches and cosmic ray hits have been removed by a
combination of outlier rejection in each spatial pixel and hand
flagging of the timestream data.  Streaking in the images results from
incorrect determinations of the zero level across scans when a bright
source is encountered and variable noise in each strip from variable
dwell time.  Atmospheric noise in observations taken in poor weather
may not be completely removed by the reduction procedures because the
atmospheric opacity can vary over timescales comparable to the time to
scan astrophysical sources (Section \ref{sec:Noise}).

\section{Flux Density and Surface Brightness Calibration}
\label{sec:FluxCalibration}

The absolute flux calibration is derived from observations of Mars and
Uranus
% no good observations, per AG
%and Neptune 
(the ``primary calibrators'').  The millimeter-wave flux of these
planets is known to $\sim 5\%$ \citep{orton86,griffin93}.
Calibrations at (sub)millimeter wavelengths are strongly affected by
atmospheric opacity corrections.  Further, the detector responsivity
of Bolocam's bolometers is a non-linear function of the mean
atmospheric loading.

To relate observations of the primary calibrators to observations of
the BGPS fields, we make use of the following relation.  The
calibration ${\cal S}$, referenced to the detectors, is given by
\begin{equation}
\label{eq:Calibration}
{\cal S} \, \left[\mathrm{\frac{V}{Jy}}\right] = 
S(\tau) \eta A \exp{(-\tau)} \Delta\nu
\end{equation}
where $S$ is the bolometer responsivity ([V/W]), $\eta$ is the system
optical efficiency, $A$ is the effective telescope collecting area,
$\Delta \nu$ is the bandwidth, and $\tau$ is the line-of-sight,
in-band atmosphere opacity.  Under the assumption that the only power
variation on the detectors is due to the power from the atmosphere
(i.e.  that astronomical sources are faint relative to the
atmosphere), which may be parameterized by $\tau$, ${\cal S}$ is a
single-valued function of $\tau$.  We have used the mean bolometer
resistance as a proxy for $\tau$, since the bolometer resistance is a
single-valued function of loading.  {\it This quantity is monitored
continuously for all observations.}  Note that this calibration curve
folds in both the effects of changing atmospheric transmission and the
changes in the detector response with optical loading.

%Multiple observations of secondary calibrators suffice to give the
%shape of the curve.  To reference it to the absolute calibrators, the
%fluxes of the planets were held fixed and the fluxes of the secondary
%calibrators allowed to float to produce the best agreement with a
%second-order polynomial for all sources.  The values for the secondary
%calibrators agree well with previously obtained Bolocam values and
%with values in the literature \citep{sandell94,jenness02}.  
%%A fit of Equation \ref{eq:Calibration} 
%

A fit to the observed values (measured potential difference vs. known
flux density) for the primary calibrators was performed for the each
epoch separately.
%CHECK
The agreement between epochs was good, so a single combined
calibration was used for all data.
%We have found that the calibration can differ
%between successive cooldowns of the instrument by $\sim10\%$ \Check\
%but is generally stable over the period of order a month which
%constitutes each of the observing epochs.  
The calibration curve is shown in Figure \ref{fig:CalibrationCurves}.
The resulting error on the calibration curve fit is less than
%CHECK
$5\%$ (statistical) over the observed range of $\tau$.  A tally of all
contributions to the flux density error is given in Table
\ref{tab:ErrorBudget}.

The above flux density calibration only accounts for the average
calibration of bolometer Volts to Janskys.  It is also necessary to
account for the variation of bolometer response across the focal
plane.  We do this by monitoring the response to the atmosphere
emission in all bolometers (astrophysical signal is negligible except
at SgrB2).  Being in the near field, the atmosphere is common to all
detectors, and thus this serves as a common relative reference to
calibrate out variation in the individual detector responsivities and
optical efficiencies.  The properties of atmospheric noise above Mauna
Kea as inferred from the Bolocam data are discussed in
\citet{sayers09b}.  The change in relative response with loading for a
typical detector is shown in Figure \ref{fig:CalibrationCurves}.  The
actual application of the method to the time series of the bolometer
data is shown in Figure \ref{fig:Flatfield}, where it is clear that
correlation of the timestreams is indeed improved by the application
of a single multiplicative factor.  The implementation of this
calibration in the context of the larger mapping algorithm is
discussed further in Section \ref{sec:Mapping}.

The instrument beam size (PSF) is measured using the planets Uranus
and Neptune which are nearly point sources for Bolocam.
%The semi-diameter of
%Mars over this period ranged from 2.3\arcsec\ to 2.9\arcsec, with
%Neptune $\sim1.1\arcsec$.  Relative to the \bcamfwhm~FWHM Bolocam
%beam, these may be regarded as point sources; the finite size of Mars
%results in a $< 1.5\%$ correction of the observed flux.
The maps of the planets, including all detectors, are fit to a
Gaussian profile.  The beam area is that averaged over all detectors,
and is $1.71\times10^{-6}$ steradians (\bcamfwhm).  This allows
conversion of the maps, calibrated in Jy beam$^{-1}$, into surface
brightness (MJy sr$^{-1}$).  The uncertainty in the mean beam size is
$3\%$, leading to a corresponding additional uncertainty in the
surface brightness calibration.

\subsection{Comparison to Other Surveys}
\label{sec:motte}

Comparison of the BGPS flux density calibration with SCUBA 850 \mum\
or ATLASGAL 870 \mum\ results is complicated by the poorly determined
variations in the dust spectral index produced by emissivity
variations.  However, there are several large-area surveys with the
IRAM 30 meter telescope at 1.2 mm which can be used for comparison
with BGPS to check overall image quality, pointing, and flux
calibration.  The study of the Cygnus-X region by \citet{motte07}
(hereafter M07) presents one of the largest fields in the Galactic
plane that has been mapped with MAMBO which overlaps with the Bolocam
survey fields.

The M07 Cygnus-X study used both the 37 element MAMBO and 117 channel
MAMBO-2 instrument on the IRAM 30-meter diameter telescope on Pico
Veleta in Spain.  The FWHM beam size for the observations was
11\arcsec .  The effective passband of the images presented by M07 is
about 70 GHz centered at 240 GHz and thus includes the bright J = 2--1
CO transition (see Figure \ref{fig:Bandpass}).

The structure of the emission in the Cygnus-X region presented by
\citet{motte07} appears nearly identical to the 1.1 mm Bolocam images
of the same field despite the inclusion of the CO line, the slightly
longer (by about 10\%) longer effective wavelength, and different beam
sizes.  The peak positions also agree to within the errors of the
BGPS, as discussed in this paper.

Direct comparison with the fluxes listed in M07 requires consideration
of the factor of three difference in the size of the beams.  For
example, the the brightest source in the Cyg-X region is DR 21.  While
M07 quote a peak flux density per beam of 4.2 \jyb for the
brightest feature in the DR 21 complex (their peak N46) and 17.9 Jy in
a $\sim$ 17\arcsec\ by 23\arcsec box (0.19 x 0.14 pc at an assumed
distance of 1.7 kpc; see M07 Table 1 in their Appendix).
BGPS measures a peak flux of 17.5 \jyb\ in an effective beam roughly
\bcamfwhm\ in diameter, indicating that the relative calibration is
similar on the brightest source in common to both surveys.  However,
DR 21 is an extended source.  Fluxes are best compared using sources
that appear compact as observed with the beam of the IRAM 30 meter
telescope.

We have convolved the \citet{motte07} fits files\footnote{{\tt
ftp://cdsarc.u-strasbg.fr/pub/cats/J/A+A/476/1243/fits/}} with a Gaussian
kernel that generates an approximately \bcamfwhm\ beam, and compared the
measured fluxes with those from the BGPS.  Additionally, the data values
in M07 are in units of Jankys per pixel, with each pixel subtending
a 4\arcsec\ by 4\arcsec\ area on the sky.  In contrast, the data values
in BGPS are in units of Janskys per beam where the beam is 31\arcsec\
in diameter.   The conversion of units required multiplication of the
convolved images by factor of 8.

The M07 images were convolved with IRAF task
GAUSS using a 3.5 pixel convolution kernel to synthesize maps with an
effective FWHM beam diameter of \bcamfwhm ;  the BGPS maps have effective
beam diameter 4\arcsec\ larger than the CSO beam diameter of 31\arcsec\
because of the smearing effects of optical distortion at the edges of the
Bolocam array.  This distortion has not been removed from the Version 1.0
images.  The Bolocam flux calibration is based on the measurement of
primary calibrators using the central pixels of the array which are not
effected by distortion.   The results are shown  in Figure
\ref{fig:CygnusComparison} and tabulated in Table 2.  This figure and
the table shows that BGPS and \citet{motte07} clumps have fluxes
consistent within about 20\% after convolution to a \bcamfwhm\ beam and
scaling to Jy/beam.

The average M07 fluxes are larger than the BGPS fluxes by a factor
of 1.23 despite the longer wavelength.  Contamination by the bright
CO J = 2-1 line may account for a few percent of the excess
detected flux by MAMBO.    For a dust spectral index of $-3.5$, the
MAMBO measurements should have detected a flux that is about 35\% lower
than what was seen by BGPS.    If this is a typical spectral index for
the sample of clumps, then the M07 flux calibration may be as
much as 45\% higher than the BGPS fluxes.

\section{Astrometry}
\label{sec:Astrometry}

\subsection{Absolute Reference Sources and Pointing Model}

We constructed an absolute reference system for the BGPS by observing
bright quasars and blazars near the Galactic Plane [with small raster
maps?].  The distribution of pointing sources over the sky is shown in
Figure \ref{fig:PointingCalibrators} and the list is given in Appendix
\ref{app:PointingCalibrators}.

Pointing observations were performed approximately once every two
hours over each night.  The elevation offsets showed a deviation from
zero which was empirically well-modeled by a quadratic function of
elevation and a linear function of azimuth.  No systematic deviation
was observed for the azimuth offsets.

% Because the Galactic Plane appears over much of the sky, the pointing
% model necessarily encompasses a large range of elevation and azimuth.
An RMS scatter of $\sim 6\arcsec$ for the model over the
entire observed range was achieved, with the scatter somewhat worse at the
highest ZA.  
Further details of the pointing calculation are given in Appendix
\ref{app:PointingCalculation}.

\subsection{Pixel Positions in the FOV}

In addition to the model for the pointing center, it is necessary to
empirically determine the actual projected pattern of the array on the
sky, and measure the offset angle between the focal plane and the sky
coordinate system.  This is done by making observations which track
all detectors across a bright source (e.g. a planet) and making maps
from each bolometer individually.

%[I think this section needs to be checked on more.... I don't know
%what to say about it yet -Adam] In our first data release, the
%projected positions of the detectors on the sky are made using the
%nominal positions without correction for optical distortion.  This
%results in a slightly blurred beam (35\farcs6 instead of 31\farcs2).
%Because this distortion is asymmetric, it can also result in a small
%pointing offset at the field edges where only bolometers with large
%offsets record data.  However, this effect is strongly mitigated by
%the low number of hits at field edges - this data is generally flagged
%out.  The residual offset is within the overall pointing model error,
%but caution should be used when measuring locations at the field
%edges. [ Note - I haven't been able to measure this yet.  It's only
%marginally possible]
%

In our first data release, the projected positions of the detectors on
the sky are made using the nominal positions without correction for
optical distortion.  This results in a slightly blurred beam
(\bcamfwhm instead of 31\farcs2).  Since a different subset of the
array is used to map each portion of the scanned field, the effective
beam size will vary slightly across the field, ranging from
$\sim31\farcs$ to $\sim33\farcs$, and it may be asymmetric.  The
effective beam size was measured from fully-sampled beam-mapping
observations of planets, while the variation in beam size with
sampling was measured from flux calibration observations of planets.
The beam size at each position cannot be measured until distortion
correction is performed in a later release.

Because this distortion is asymmetric, it can also result in a small
pointing offset at the field edges where only bolometers with large
offsets record data.  However, this effect is strongly mitigated by
the low number of hits at field edges - this data is generally flagged
out.  The residual offset is within the overall pointing model error,
but caution should be used when measuring source locations at the
field edges.

Because this distortion is asymmetric, it can also result in a small pointing
offset at the field edges where only bolometers with large offsets record data.
However, this effect is strongly mitigated by the low number of hits at field
edges - this data is generally flagged out.  The residual offset is within the
overall pointing model error, but caution should be used when measuring
source locations at the field edges.

A mean focus offset for the secondary was determined once per
observing run.  An elevation-dependent correction is applied via a
CSO-determined look-up table between raster scans.  We did not see any
evidence that more frequent focus checks improve the quality of the
PSF.
%[Observers - what checks were done?]

\subsection{Relative Alignment and Mosaicking}

Relative alignment between different observations of the same field
was performed by finding the peak of the cross-correlation between
images and a pointing master constructed from observations taken in
the epoch with the best-constrained pointing model for that field (in
most cases, this was from Epoch V, though in some cases Epoch VI was
used).  Each observation was initially mapped individually, then all
observations in the pointing epoch were median-combined to create a
reference field.  Each observation was then cross-correlated with the
reference image for that field. The cross-correlation peak was fit
with a gaussian and the difference between the gaussian peak and the
image center was used as the pixel offset.  Finally, all observations
of a field were merged into a single timestream with pointing offsets
applied to create the field mosaic.

This method of alignment avoids the ambiguities inherent in using
extracted sources to align fields, as the BGPS sources are rarely
point-like.  It further avoids the pointing smearing and loss of peak
flux density which would result if the the maps of individual
observations were coadded using the pointing model alone, where each
individual observation would be coadded with the 6\arcsec\ RMS
uncertainty of the model.  The relative alignment using the
correlation maps may be determined accurately to the error in finding
the peak of the gaussian, typically $<1\arcsec$ where there is
adequate signal.  In a few low S/N fields, specifically \lon=65 to 75,
there was not enough signal to acquire a pointing offset using
cross-correlation.  In these fields, the pointing model was used
directly.  This results in a slightly larger effective beam since the
pointing error of 6\arcsec\ is convolved with the instrument PSF.

\subsection{Comparison to SCUBA Legacy Catalog Source Positions}
\label{sec:SCUBAPointingComparison}
%and SHARC-II Positions}

All SCUBA 850 and 450 \mum\ data has been re-processed in a uniform
manner and made publicly available as the SCUBA Legacy Catalog
\citep[][hereafter SLC]{difrancesco08}.  These maps allow a
cross-check of the accuracy of the BGPS pointing model.  We have
applied the same cross-correlation procedure used to obtain the
relative alignment of BGPS observations to compare the BGPS images to
the SLC.  Concerns might arise in comparing the Bolocam to the SCUBA
maps, such as the differing scan strategies (SCUBA maps were typically
taken in jiggle-map mode, which tends to remove extended structure).
The cross-correlation technique actually mitigates these effects,
since it does not require the establishment of a single position for
an object (which may differ due to the change in wavelength between
the surveys or due to the observing strategy) but rather makes use of
all the information available.  The by-eye morphological comparison
between SCUBA and BGPS sources is generally excellent with some
exceptions, giving a high degree of confidence to the comparison and
to the results obtained from the BGPS pointing model [include Di
Francesco pointing accuracy].  An example is shown in Figure
\ref{fig:SCUBAPointingComparison}.  We find that the RMS offset
between BGPS and SLC positions within a $3\arcdeg \times 1\arcdeg$
block defined by the BGPS is $4.0\pm8.8$\arcsec, consistent with the
errors derived for the Bolocam pointing model.  There were some BGPS
fields that had larger offsets from SCUBA fields; these fields were
excluded from the offset measurement above because of poor
morpohological match.

For the purposes of future comparison of both flux and pointing with
other surveys, we present in Appendix \ref{app:ReferenceSources} a
selection of bright, compact, relatively isolated sources.

%{\em Miranda's data here}

\section{Mapping Algorithm}
\label{sec:Mapping}

The essential signal-processing problem to be solved by the mapping
algorithm is the estimation and removal of a signal due to fluctuating
atmospheric emission which is $>\sim$100's of times stronger than the typical
sources in our maps.  We implement an iterative procedure for
estimating both the atmosphere fluctuations and the astrophysical
signal without {\it a priori} knowledge of either.

We assume the raw timestream data $d$ for each bolometer (indexed by
$i$) can be written as
\begin{equation}
d_i(t) = s_i(t)+a_i(t)+e_i(t)+\epsi_i(t)
\end{equation}
where $s$ is the astrophysical signal, $a$ is atmospheric fluctuation
noise, $e$ are non-random signals due to the instrument itself, and
$\epsi$ is irreducible Gaussian noise due to photon shot and detector
noise.  The process of making a maximum likelihood map in the presence
of Gaussian noise is well-understood, and if the only contributions to
the data were $s$ and $\epsi$, then the minimum variance estimator of
the true astrophysical map is given by
\begin{equation}
\label{eq:LSMap}
m = (A^T W A) A^T W d
\end{equation}
where $A$ encodes the pointing information and $W$ is the inverse of
the covariance matrix of the time domain noise $\epsi$,
\begin{equation}
W = {\langle \epsi_i \epsi_j \rangle}^{-1}
\end{equation}
In this case, the mapping from data to map is a linear operator, $M
=(A^T W A) A^T W$.  For compactness, Equation \ref{eq:LSMap} will be
written as $M[d] = m$.  Note that the mapping operation is not
invertible.  However, given a map $m$ and an observing matrix $A$,
there exists a linear operation which makes predictions about the
observed timestreams, namely $d = A m$; this will be denoted $T[m] =
d$. The goal is to produce a time series for each bolometer which as
closely as possible approximates $s + \epsi$ so that we may produce
the best estimate of the astrophysical signal $m = M[s+\epsi] = S +
N$.

We proceed iteratively, estimating $a$ and $e$ in the order of their
relative strengths, beginning with the atmosphere.  The first
iteration step is then
\begin{equation}
\label{eq:FirstIterStep}
\tilde{a}^{(n)}(t) = 
\sum_i{d_i(t)-\tilde{s}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)}
\end{equation}
where $\tilde{s}^{(0)}_i(t)=0$.   Each bolometer's timestream is then 
fit to the mean atmosphere model 
\begin{equation}
\label{eq:MeanAtmosphere}
\tilde{a}^{(n)}_i(t) = \frac{d_i(t) \cdot a(t)}{d_i(t)^2} a(t) = r_i a(t)
\end{equation}
This is illustrated in Figure \ref{fig:Flatfield}.  The $r_i$ are then
used as the relative gains of the detectors.
% James - what's this section about?
% The instrument
% errors, which are typically smaller than the atmosphere, are then
% estimated
% \begin{equation}
% \tilde{e}^{(n)}_i(t) = some models
% \end{equation}
Both the atmosphere and instrument models are subtracted from the
raw time series
\begin{equation}
\label{eq:SignalEstimate}
d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)
\approx s_i(t) + \epsi_i(t)
\end{equation}
This is the best estimate of signal plus irreducible noise at
iteration step $n$, and is made into a map
\begin{equation}
M[d_i(t) - \tilde{a}^{(n)}_i(t) - \tilde{e}^{(n)}_i(t)] = \tilde{m}^{(n)}
\end{equation}
The current best map $\tilde{m}^{(n)}$ is then deconvolved to provide
a relatively low-noise, smooth map from which to generate a timestream
\begin{equation}
\label{eq:MapToTime}
T[{\cal D}[\tilde{m}^{(n)}]] = \tilde{s}^{(n)}_i(t)
\end{equation}
where ${\cal D}$ represents the deconvolution operation.  At this
stage, the iterative process could begin again with Equation
\ref{eq:FirstIterStep}, though it is useful to produce a
``residual map''
\begin{equation}
\label{eq:ResidualMap}
E = M[d_i(t) - \tilde{s}^{(n)}_i(t) - \tilde{a}^{(n)}_i(t) - 
\tilde{e}^{(n)}_i(t)]
\end{equation}
This residual map serves as a visualization of the progress of the
iterative method, since in a perfect process it would be map of
$\epsi$.

\subsection{Atmosphere Fluctuation Noise Model}

By construction, the Bolocam beams as they pass through the dominant
layer of atmospheric water vapor at $\sim1$~km altitude are still
highly overlapped, and therefore sample a nearly identical region of
atmosphere.  The simplest model for the atmosphere fluctations makes
use of this fact by assuming that the largest correlated signals
between detectors are due to to the atmosphere correlations.  However,
the simple common mode subtraction of Equation \ref{eq:MeanAtmosphere}
does not remove all of the correlated signal.  To further remove the
atmosphere (and other correlated instrument noise) without a detailed
physical model, we use principal component analysis (PCA) as given in
\citet{laurent05}.
%[more words]
While lacking the physical motivation of the polynomial subtraction
used for the 2.1 mm Bolocam data by \citet{sayers09b}, the 1.1 mm data
seems to have a large number of correlated components that are not
simply described by low-order polynomials in the focal plane.  Figure
\ref{fig:PCA_Graphical} shows an example of the modes removed by the
PCA model.  In addition to the PCA model, after iteratively
subtracting the astrophysical model, each scan has a polynomial fit in
time removed to deal with the longest time scale modes.

%The mean across each scan is forced to zero by subtracting a
%polynomial fit to the scan after iteratively rejecting astrophysical
%sources.  This process helps reduce negative bowls in the scan
%direction near bright sources.

\subsection{Instrument Error Signals}

Most of the instrument error signals have characteristic features
which allow them to be identified, though not always removed.  The
error signals include the following:
\begin{enumerate}

\item Pickup from the 60 Hz AC power.  This appears as narrow lines in
    the power spectral density (PSD) of the data.  The second harmonic
    of 60 Hz is aliased via beating against the 130 Hz bolometer bias
    frequency into 10 Hz with sidebands split at $\sim1$ Hz.  It is
    removed by replacing components at these frequencies in the
    fourier transform with gaussian noise matched to the local mean
    amplitude.

\item Spikes in voltage due to cosmic ray strikes on the bolometers
(``glitches'').

% James - technical question: Why is there microphonic pickup?  What
% does the term mean, exactly?
\item Microphonic pickup due to vibrations of the receiver.  The most
noticeable microphonics occur at the end of each scan when the
telescope is turning around and the field rotator is adjusting.  This
leads to broad spikes in the time series, whose long decay must be
removed from the data, particularly during the beginnings of scans.

%\item Others?

\end{enumerate}

Fortunately, most of the AC powerline pick-up occurs at frequencies
where there is no astrophysical signal, given the beam size and scan
speed.  Thus this error is dealt with by first notch filtering at the
line frequencies and then low-pass filtering the data.  Because of the
low correlation with astrophysical signal, this step is performed only
once and is not iterated (c.f. Figure 1 of \citet{sayers09}).

Both glitches and microphonic pickup from the scan turnarounds are
degenerate with astrophysical signal and must be estimated as part of
the iterative process.  
% NOT TRUE - they are identified per-pixel most of the time
% Glitches are identified as large excursions
% from the RMS level after subtraction of the best atmosphere and bright
% source model, and the data there is excluded from inclusion in
% subsequent maps.
The turnaround microphonics are modeled as decaying exponentials at the
beginnings and ends of scans and are also removed by the PCA cleaning.

\subsection{Data Flagging}

Due to the large volume of data generated by the survey, it was
necessary to develop new tools to quickly visualize the data and
ensure data quality.  A common tool in radio astronomy is the
so-called ``waterfall'' plot, which is an image with frequency on the
x-axis, time on the y, and an intensity proportional to the
interferometer visibility amplitude or phase.  We have used a variant
on this to visualize the Bolocam data by displaying bolometer number
(related to the bolometer position in the focal plane) on the x-axis
and time on the y, with the intensity given by the bolometer's
response (in Jy) at that time.  An example of this is shown in Figure
\ref{fig:Flagger}.  Anomalies such as the glitch shown
can be detected and manually and interactively flagged out.

An automated flagger was also created that flags out outlier data on a
per-map-pixel basis.  In order to make a robust measurement of the
variance of the fluxes assigned to each pixel, we used the median
average deviation (MAD) over the data in the pixel and rejected high and low
outliers at the 3-sigma level.  Pixels with too little data to compute
a deviation, i.e. those with $<3$ data points, were also flagged out -
these scan-edge pixels are the dominant contribution to the total
number of flagged data points.

\subsection{Creation of the Astrophysical Model}

The timestream data is made into a spatial map using the pointing data
corresponding to each time point for each bolometer.  The data is
weighted by inverse variance across a single scan and then drizzled
into a map with $7.2$\arcsec\ pixels using a nearest-neighbor
algorithm.  The nearest-neighbor matching allows the map to be
returned to a timestream in the same manner, but with the S/N improved
by averaging over all hits on a given pixel.  It has the disadvantage
that it accentuates the unevenness of the cross-scan sampling, since
each timestream point's value is assigned to an area much smaller than
the beam.  In principle this can be addressed by various gridding
algorithms, and may be a part of future reductions of the BGPS data.

The resulting map is then subjected to a maximum-entropy clean
algorithm \citep{hollis92} using a specified kernel to produce a
deconvolved image which is mapped back into the timestream to be
subtracted (Equation \ref{eq:MapToTime}).  This has several advantages
over using the original map directly, including rejecting artifacts
smaller than the kernel and decreasing the noise of the resulting
timestream.  It is also better than a ``threshold'' method, which
tends to produce discontinuities and include noise outliers.

In some fields, flux from some sources spread out over the course of
the iterative process using certain kernel sizes.  The cause of this
artifact is not yet well-understood, but using a different kernel size
was an effective workaround and did little to change the properties of
the final map.  The standard kernel size was 14.4\arcsec; a
21.6\arcsec kernel was used when artifacts were encountered.
Simulations indicate that this change in kernel has a negligible
effect on the flux of the source; see Figure \ref{fig:PCA_Filter}.

\section{Noise and Systematic Effects}
\label{sec:Noise}

The residual maps $E$ (Equation \ref{eq:ResidualMap}) form the basis
for the noise estimation and also provide a way of estimating the
systematic error resulting from imperfect subtraction of bright
sources.  These maps are produced for each region from the data
residuals after removing the astrophysical signal model and the PCA
atmosphere model.  An example of an error map is shown in Figure
\ref{fig:sampleimages}.  In an ideal case, the residual map $E$
represents a map of a realization of the underlying irreducible noise
$\epsi$.  However, the imperfect estimation of the signal results in
positive ``ghosts'' of bright sources.  We have found that smoothed
versions of the residual maps provide the best means of determining
the pixel-to-pixel error that account for the local variations in the
noise, including those due to such artifacts.
%Near the scan edges,
%1/sqrt(weightmap) provides a better estimate of the local noise.
% ...ok, explain please?
Our pixel noise estimate assumes the noise is uncorrelated between
adjacent pixels, but of course adjacent spatial pixels are correlated
due to residual atmospheric noise correlated in time.  We have ignored
this effect in our analysis under the assumption that the remaining
correlations after subtracting the PCA model are small.

Because observing conditions varied widely during the survey, the RMS
noise levels obtained in the various fields of the BGPS also varies.
However, the noise within a given $1\arcdeg \times 1\arcdeg$ field is
fairly uniform in the absence of very bright sources.  We show the
variation of the depth as a function of Galactic longitude in Figure
\ref{fig:NoiseVsLongitude}.

\section{Effects of the Data Reduction Procedures on Photometry}
\label{sec:Photometry}

To do accurate and meaningful photometry on the Bolocam maps requires
a good estimate of the noise (Section \ref{sec:Noise}) but also an
understanding of the spatial filtering imposed by the observing
strategy and the cleaning and mapping of the data.  The additional
complexities of defining ``sources'' in complicated regions are
discussed in \citet{rosolowsky09}.  Here we restrict our discussion to
the limitations on performing photometry imposed by the data
reduction.

The fundamental feature of the data is that there is a degeneracy
between large spatial scale astrophysical emission and the atmosphere
and other signals with long spatial or temporal variations.
Consequently, there are limits to ability of the algorithm to separate
the signal $s$ from the atmosphere $a$ and instrument terms $e$ in
Equation \ref{eq:SignalEstimate}.  Inevitably, some of the signal
which should be present in $s$ is mixed into these terms and is thus
not present in the final map $m$.  Because $s$ is generated from a
deconvolution algorithm, it has non-zero mean by construction, but
importantly, it does {\it not} contain an unbiased estimate of the
astrophysical signal on all spatial scales: large spatial scales are
preferentially missing.

To assess the effect of the mapping algorithm on the flux recovered as
a function of angular scale, we performed a series of simulations.
Gaussian sources with a range of peak flux densities and sizes were
inserted on top of a background consisting of real BGPS data with the
astrophysical source model removed.  The recovered fluxes were
measured in ellipses that included single well-separated sources.
Figure \ref{fig:PCA_Filter}a shows the fraction of flux recovered in
the map as a function of source size for a range of different PCA
components subtracted in the cleaning.

The final reduction used 13 PCA components because this produced a
reasonable compromise between attenuating extended structure and
cleaning the atmospheric contribution.  Note that this is a
substantially more aggressive cleaning than that used in
\citet{enoch06} largely because our mean integration time per pixel
was $\sim$seconds, whereas theirs was $\sim$minutes.  
%, and therefore the
%atmospheric noise was much more likely to be correlated between
%observations.

Figure \ref{fig:PCA_Filter}b shows the fraction of flux recovered in
the maps as a function of the deconvolution kernel used to create the
astrophysical model maps.  The differences at large source sizes are
not well-understood (these size scales may be dominated by
atmosphere), but at the smallest source size, it is evident that a
deconvolution kernel smaller than the source size does not reliably
recover the source flux.  Therefore we chose a deconvolution kernel
smaller than the beam but larger than the pixel size used.  Figure
\ref{fig:Deconvolution} demonstrates the effects of using both too
large and too small a kernel.

\section{Final Maps and Data Release}
\label{sec:FinalMaps}

The final maps are produced by coadding all observations of a single
square degree centered on a particular Galactic coordinate
$(l_0,b_0)$.  The maps are made in Galactic coordinates using a plate
carr\'{e} (FITS header CAR) projection.  (This is the same projection
used by the \spitzer-GLIMPSE and BU-FCRAO GRS surveys.)  As these maps
are near the coordinate system equator, the difference between truly
equiareal pixels and the pixels used is at most even at 5\arcdeg out
of the plane (as for IC1396) is 0.4\%.  The pixel size is 7.2\arcsec,
chosen to be much smaller than the Bolocam beam.  The maps are written
to standard FITS files.  An example of the FITS header for the BGPS
images is given in Appendix \ref{app:FITS_Header}.

Five image types will be released, each covering the same regions.
The maps are standard astrophysical maps subject to the caveats
previously mentioned, the most important being the spatial filtering
function (section \ref{sec:Noise}).

All processed maps are available through
IPAC\footnote{http://irsa.ipac.caltech.edu/Missions/bolocam.html}.
IPAC provides a cutout service for the images as well as a searchable
version of the catalog provided in \citet{rosolowsky09}.  The full
release includes:

\begin{enumerate}
\item Calibrated maps for all regions observed (MAP).

% don't release smoothmaps? \item The MAP smoothed by the Bolocam PSF, which is helpful in bringing out extended structure (SMOOTHMAP)

\item The deconvolved MAP (MODEL).  The model maps are deconvolved
using the maximum-entropy method with kernels specified in the header
(either 14.4\arcsec or 21.6\arcsec).  {\tt After deconvolution, the maps
are smoothed to the beam size of \bcamfwhm.}

\item Maps of integration time per pixel (NHITSMAP).  The nhitsmap
shows how many recorded data points from the timestream have been
assigned to each pixel; it is proportional to the dwell time per pixel
(each `hit' is .1s).

\item Maps of residual noise per pixel (RESIDUAL).  The residual map
is the map of the model timestream subtracted from the data
timestream; it is close to but not identical to the map minus the
model.

\item Maps of inverse variance per pixel (WEIGHTMAP).  The weightmap
shows the inverse-variance weighting of each pixel.

\item Maps of the mask obtained for various regions from the
cataloging algorithm (MASKMAP) \citep{rosolowsky09}

\end{enumerate}

Examples of each (except the nhitsmap) are given in figure
\ref{fig:sampleimages}.

\section{Discussion}
\label{sec:Discussion}

%{\bf Claudia suggested that this section be removed or changed to an
%`advertisement' section where we indicate what follow-up work needs to
%be done - seems reasonable, since this is mostly junk pilfered from
%the `letter'.  However, we need to keep the last sentence around...}

The BGPS is the first millimeter survey of a substantial fraction of
the Galactic Plane, providing an unbiased look at the high-density gas
most intimately associated with the earliest phases of star formation.
Maps of the entire BGPS are presented in Figure \ref{fig:BGPSMontage}.

Inspection of Figure \ref{fig:BGPSMontage} shows excellent
correspondence with SCUBA maps of small sub-fields throughout the
plane \citep{difrancesco08}.  BGPS contours overlaid on the GLIMPSE
survey \citep{benjamin03} show excellent correlation with 8\mum\
infrared dark clouds (IRDCs).  However, many features in the BGPS maps
do not appear in the Spitzer observations, which suggests that the 1.1
mm emission traces dense clumps without a background to absorb; see
Figure \ref{fig:Science}.

% No advertisements.
% A
% study of IRDC correlation with BGPS emission is underway
% \citep{battersby09}.

% implies star forming....

Filamentary structures with aspect ratios $\geq 10$ are present
throughout the galactic plane.  The most crowded fields consist of
frothy and clumpy structure.  There are 5000-10000 distinct sources,
the majority of which are at least moderately resolved
\citep{rosolowsky09}.  We suspect that higher resolution observations
will resolve BGPS clumps into clusters of protostellar objects, as has
been seen in Serpens \citep{enoch08,Testi1998}, OMC 1
\citep{beuther04,johnstone99}, and S255N \citep{cyganowski07} where
bolometer observations at low resolution have been complemented by
interferometer observations at high resolution.

There are clear maxima in the emission filling factor in the Galactic
Center and in the region from $\lon=23$ to $\lon=31$, the line of
sight through the densest part of the molecular ring.  Additional
areas of high source density are seen at $\lon=10$ and $\lon=13$.  The
filling factor of 1.1 mm continuum emission at the 100 mJy level is
$\sim 20$\% per square degree from $\lon=-10$ to $\lon=20$ and $\sim
10$\% from $\lon=20$ to $\lon=50$ (compare to figure
\ref{fig:NoiseVsLongitude}.  In the range $\lon=50$ to $\lon=75$, the
filling factor is below 1\% and very few sources have been detected.
This result is in contrast to CO filling factors $\sim1$ in the inner
galactic plane \citep{dame01,jackson06}, confirming that 1.1 mm
continuum emission traces denser material than CO.  The small filling
factor of the 1.1 mm emission compared to CO emission shows that the
continuum is directly associated with star forming material, whereas
the CO also traces less dense material not directly associated with
star formation.

In some cases, distance to BGPS clumps can be determined by matching
Galactic Ring Survey $^{13}$CO and 1.1 mm continuum morphology
\citep{IRDCdistance,jackson06}.  However, in other cases the
association between the CO and 1.1 mm data is not clear, either
because of confusion or because the BGPS sources are too compact to
identify in $^{13}$CO morphology.  Heterodyne follow-up observations
using dense gas tracers (NH$_3$, N$_2$H$^+$, HCO$^+$, and CS) are
being conducted to provide radial velocity, density, chemistry, and
temperature measurements \citep{schlingman09,battersby09, nordhaus09}.

The depth can be converted into an estimate of mass via standard
estimates, i.e., under the assumption of a temperature, opacity, and
distance.  Using dust opacity from \citet{ossenkopf94}
%\citet{enoch06} and \citet{young06},
and assuming a gas-to-dust mass ratio 100, the mass sensitivity
can be written as
\begin{equation} 
\label{eq:Mass}
M_{gas}\approx
0.22\left( e^{13.1/T_d}-1 \right) \left({S_\nu\over 15\; {\rm mJy}} \right)
D^{2}_{kpc} \msol 
\end{equation}
where $D_{kpc}$ is the distance to the source in kpc.  (At $S_\nu =
15$~mJy this corresponds to an RMS extinction sensitivity of $A_V\sim
1$ mag, with the same $T_d$ dependence).  For Bolocam, $h\nu/k \approx
13$~K, so for $T_d > 10$K the dependence is approximately linear.  At
the coldest temperatures expected, the mass changes a factor of 2.9
between 20 and 10 K.  The distance to the clump is more important in most
situations.

BGPS clumps are associated with the densest portions of giant
molecular clouds (GMCs) traced by CO surveys
\citep[e.g.,][]{dame01,jackson06}.  The brightest clumps visible in
our maps, e.g. \object{Sgr B2}, \object{G34.3+0.15}, \object{W 51},
\object{W 43}, \object{W 49}, and \object{M 17}, correspond to well
known star forming regions.
%(the saturation is only a display artifact; the astrophysical sources are
%always much fainter than the atmosphere and therefore do not saturate the
%detectors).  
These associations suggest that other bright sources of millimeter
emission will prove to be massive star or cluster forming regions.

We anticipate that the public release of the BGPS data will serve as
an important first look to help guide HIGAL, JPS, ...

\noindent---------------------------------

{\it Facilities:} \facility{CSO (Bolocam)}

\acknowledgments

We would like to acknowledge the staff and day crew of the CSO for their
assistance. The BGPS project is supported by the National Science Foundation
through NSF grant AST-0708403. J.A. was supported by a Jansky Fellowship from
the National Radio Astronomy Observatory (NRAO). The first observing runs for
BGPS were supported by travel funds provided by NRAO. Support for the
development of Bolocam was provided by NSF grants AST-9980846 and AST-0206158.
Team support was provided in part by NSF grant AST-0607793 to the
University of Texas at Austin.

We recognize and acknowledge the cultural role and reverence that the
summit of Mauna Kea has within the Hawaiian community. We are
fortunate to conduct observations from this mountain.

\appendix

\section{Calculation of Color Corrections}
\label{app:ColorCorrections}

If an experiment has finite bandwidth ($t(\nu) \ne \delta(\nu -
\nu_c)$), to report a source surface brightness at a single
frequency, one must assume a source spectrum.  The power detected from
that source is assumed to be
\begin{equation}
P_{in} = \eta \; A \Omega \; \int I_0(\nu) t(\nu) d \nu
\end{equation}
(Here $\eta$ and $A \Omega$ are the optical efficiency and throughput
of the instrument, $I_0(\nu)$ is the nominal (assumed) surface
brightness of the source, and $t(\nu)$ is the bandpass transmission
normalized to $1.0$ at its peak.)  The effective band center $\nu_c$
is usually chosen such that
\begin{equation}
I_0(\nu_c) \simeq \frac{\int I_0(\nu) t(\nu) d \nu}
{\int t(\nu) d \nu} 
\end{equation}
The band centers for Bolocam are calculated assuming a Rayleigh-Jeans
(RJ) source spectrum
\begin{eqnarray}
I_0(\nu_c) & = & I_{RJ}(\nu_c) \\
\nonumber & = & \tau \; 2 k T \; \frac{\nu_c^2}{c^2},
\end{eqnarray}
where $\tau$ is the optical depth of the source and $k$ is Boltzmann's
constant.  Since the detected power is assumed to be
\begin{equation}
P_{in} = \eta \; A \Omega \; \int \tau \; 2 k T \; \frac{\nu^2}{c^2}
\; t(\nu) d \nu,
\end{equation}
we can write
\begin{equation}
I_0(\nu_c) = \frac{P_{in}}{\eta \; A \Omega \int \nu^2 \; t(\nu) d \nu} \; \nu_c^2
\end{equation}
Now if we assume a different source spectrum, for example a greybody
with power-law emissivity, the assumed input power is
\begin{eqnarray}
P_{in} & = & \eta \; A \Omega \; \int I_{GB}(\nu) t(\nu) d \nu \\
\nonumber & = & \eta \; A \Omega \; \int \tau(\nu_0) (\nu / \nu_0)^\alpha B_\nu(T) t(\nu) d \nu
\end{eqnarray}
and the source spectrum inferred from the detected power is
\begin{eqnarray}
I_{GB}(\nu_c) & = & \tau(\nu_0) (\nu_c / \nu_0)^\alpha B_{\nu_c}(T) \\
\nonumber & = & \frac{P_{in}}{\eta \; A \Omega \int \nu^\alpha B_\nu(T) t(\nu) d \nu} \; \nu_c^\alpha B_{\nu_c}(T) \\
\nonumber & = & \frac{\int \nu^2 t(\nu) d \nu}{\int \nu^\alpha B_\nu(T) t(\nu) d \nu} \; \nu_c^{\alpha-2} B_{\nu_c}(T) \; I_{RJ}(\nu_c) \\
\nonumber & \equiv & \frac{I_{RJ}(\nu_c)}{K}.
\end{eqnarray}
This defines the color correction $K$ to apply to the reported Bolocam
flux from a source if the source is assumed to have a greybody
spectrum with power-law emissivity.

%We make a similar calculation for DIRBE, for which the band centers are 
%computed assuming a spectrum with $\nu I(\nu)$ constant.  In this case, 
%the correction is given by 
%\begin{eqnarray}
%K &=& \frac{\nu_c^{-1}}{\int \nu^{-1} t(\nu) d \nu} \left[\frac{\tau(\nu_0) (\nu_c / \nu_0)^\alpha B_{\nu_c}(T)}{\int \tau(\nu_0) (\nu / \nu_0)^\alpha B_{\nu}(T) t(\nu) d \nu} \right]^{-1} \\
%\nonumber &=& \frac{\int \nu^\alpha B_{\nu}(T) t(\nu) d \nu}{\int \nu^{-1} t(\nu) d \nu} \; \nu_c^{-(\alpha+1)} B_{\nu_c}(T)
%\end{eqnarray}
%
We note that for an arbitrary experiment with bandpass $t(\nu)$ that
reports its surface brightness measurements assuming a spectrum
$I_0(\nu)$, the surface brightness assuming a different source
spectrum $I_1(\nu)$ is given by
\begin{eqnarray}
I_1(\nu_c) &=& I_0(\nu_c) / K \\
\nonumber &=& I_0(\nu_c) \frac{I_1(\nu_c)}{\int I_1(\nu) t(\nu) d \nu} \left[\frac{I_0(\nu_c)}{\int I_0(\nu) t(\nu) d \nu} \right]^{-1}
\end{eqnarray}

% James - it would be nice to include some numbers here, not just the process.

\Table
{llll}
{Comparison of band centers and flux ratios for Bolocam and MAMBO.}
{         & MAMBO         & Bolocam       & $S_{\rm BOLOCAM}/$ \\
 $\alpha$ & $\nu_c$ [GHz] & $\nu_c$ [GHz] & $S_{\rm MAMBO}$ }
{tab:ColorCorrections}
{
1.0 & 254.8 & 267.8 & 1.051 \\
1.5 & 256.4 & 268.6 & 1.072 \\
2.0 & 258.0 & 269.4 & 1.090 \\
2.5 & 259.5 & 270.0 & 1.104 \\
3.0 & 261.1 & 270.7 & 1.115 \\
3.5 & 262.5 & 271.4 & 1.124 \\
4.0 & 264.0 & 272.2 & 1.130 \\
4.5 & 265.4 & 272.9 & 1.134 \\
5.0 & 266.8 & 273.8 & 1.137 \\
}

\section{Pointing Calculation}
\label{app:PointingCalculation}

To calculate the position at which to point the telescope, the CSO
antenna computer computes an alt/az ($E$,$A$) derived from the RA/Dec J2000
(heliocentric) catalog position.  It performs the following
calculation:

\begin{enumerate}

\item Precess catalog coordinates to current epoch.  The CSO by
default stores catalog positions in B1950.

\item Add aberration and nutation corrections.  This is the
``requested apparent'' RA and Dec reported by the antenna computer.

\item The equatorial coordinates are transformed to horizon
coordinates using the current local apparent sidereal time.

\item An 11-term model (``$C$-terms'') based on the optical pointing of
the telescope is then applied to $A$ and $E$ along with $t$-terms (?), %James?
the tilt of the alidade, and a refraction correction to obtain the
necessary mechanical pointing of the telescope to acquire the source.

\item The telescope then attempts to follow this track; the difference
between the requested $A$,$Z$ and the actual, measured angles of the
encoders (a result of servo errors) are also reported.

% are c4 and c5 different from the alidade_x,y tilt that are in the RPC file?

%char	32   pointing_file_name
%double	11   pointing_constant
%double	6    t_term_constant
%double	1    alidade_x_tilt
%double	1    alidade_y_tilt
%double	1    azimuth_t_term_offset
%double	1    elevation_t_term_offset
%double	1    tropospheric_refraction

\end{enumerate}

The CSO antenna computer reports this pointing information, including
geocentric RA/Dec, at a rate of 100 Hz.  This time series, along with
the field rotator encoder information from Bolocam is merged with the
bolometer time series and aligned.

To calculate the pointing used in the map from the telescope inputs,
the Bolocam software does the following:

\begin{enumerate}

\item Start with the reported $\alpha',\delta'$ from the CSO.  Remove
aberration and nutation corrections.  Precess from current epoch to
J2000.  This gives $\alpha,\delta$.

\item Apply the $\Delta A,\Delta E$ terms to $\alpha,\delta$ to
account for the difference between commanded and actual positions of
the telescope.

\end{enumerate}

\section{Table of Absolute Pointing Sources}
\label{app:PointingCalibrators}

The absolute reference frame for the BGPS pointing model was chosen
from a selection of bright quasars near the Galactic Plane.  Table
\ref{tab:PointingCalibrators} gives the pointing sources used for this
purpose.

\Table
{llrrrrc}
{Absolute Pointing Calibrators}
{Number & Alias & RA(J2000) & Dec(J2000) & l & b  & Flux (Jy)}
{tab:PointingCalibrators}
{
         1 &   1622-253 & 16 25 46.9 & -25 27 38.3 & 352.14 &  16.32 &  0.18 \\
         2 & 16293-2422 & 16 32 22.9 & -24 28 35.6 & 353.94 &  15.84 &  8.40 \\
         3 &   1657-261 & 17 00 53.2 & -26 10 51.7 & 356.70 &   9.75 &  0.07 \\
         4 &   NGC6334I & 17 20 53.4 & -35 47  1.7 & 351.42 &   0.64 &  2.50 \\
         5 &    NRAO530 & 17 33 02.7 & -13 04 49.5 &  12.03 &  10.81 &  2.70 \\
         6 &   1741-038 & 17 43 58.8 &  -3 50  4.6 &  21.59 &  13.13 &  1.44 \\
         7 &   1749+096 & 17 51 32.8 &   9 39  0.7 &  34.92 &  17.65 &  2.70 \\
         8 &      G5.89 & 18 00 30.4 & -24 04  0.5 &   5.89 &  -0.39 &  3.57 \\
         9 &        M8E & 18 04 53.0 & -24 26 39.4 &   6.05 &  -1.45 &  3.92 \\
        10 &     G10.62 & 18 10 28.7 & -19 55 49.8 &  10.62 &  -0.38 & 17.20 \\
        11 &   1830-211 & 18 33 39.9 & -21 03 40.0 &  12.17 &  -5.71 &  1.70 \\
        12 &      G34.3 & 18 53 18.6 &   1 14 58.3 &  34.26 &   0.15 & 31.30 \\
        13 &   1908-202 & 19 11 09.6 & -20 06 55.0 &  16.87 & -13.22 &  1.00 \\
        14 &      G45.1 & 19 13 22.1 &  10 50 53.4 &  45.07 &   0.13 &  3.70 \\
        15 &      OV236 & 19 24 51.0 & -29 14 29.8 &   9.34 & -19.61 &  8.60 \\
        16 &   1923+210 & 19 25 59.6 &  21 06 26.1 &  55.56 &   2.26 &  0.32 \\
        17 &   1954+513 & 19 55 42.7 &  51 31 48.6 &  85.30 &  11.76 &  0.60 \\
        18 &       CYGA & 19 59 28.5 &  40 44  1.7 &  76.19 &   5.75 &  0.25 \\
        19 &      K3-50 & 20 01 45.7 &  33 32 43.5 &  70.29 &   1.60 &  8.45 \\
        20 &   2005+403 & 20 07 44.9 &  40 29 48.6 &  76.82 &   4.30 &  0.40 \\
        21 &        ON1 & 20 10 09.1 &  31 31 37.7 &  69.54 &  -0.98 &  2.90 \\
        22 &    2013+37 & 20 15 28.7 &  37 10 59.6 &  74.87 &   1.22 &  1.40 \\
        23 &   2021+317 & 20 23 19.0 &  31 53  2.4 &  71.40 &  -3.10 &  0.54 \\
        24 &   2023+336 & 20 25 10.8 &  33 43  0.3 &  73.13 &  -2.37 &  0.92 \\
        25 &     GL2591 & 20 29 24.7 &  40 11 18.9 &  78.89 &   0.71 &  2.07 \\
        26 &    MWC 349 & 20 32 45.5 &  40 39 36.7 &  79.64 &   0.47 &  1.67 \\
        27 &       W75N & 20 38 36.4 &  42 37 34.5 &  81.87 &   0.78 &  4.46 \\
        28 &      3C418 & 20 38 37.0 &  51 19 12.7 &  88.81 &   6.04 &  0.62 \\
        29 &   CRL 2688 & 21 02 18.8 &  36 41 37.7 &  80.17 &  -6.50 &  1.90 \\
        30 &    NGC7027 & 21 07 01.6 &  42 14 10.2 &  84.93 &  -3.50 &  9.27 \\
        31 &   2201+315 & 22 03 15.0 &  31 45 38.4 &  85.96 & -18.78 &  0.35 \\
}

\clearpage

\section{Table of Reference Sources}
\label{app:ReferenceSources}

\Table
{llrrrrc}
{In-plane Millimeter Sources}
{Number & Alias & RA(J2000) & Dec(J2000) & l & b  & Flux (mJy)}
{tab:PPS}
{
   l351pps &            & 17 23 50.2 & -36 38 58.8 & 351.04 &  -0.34 &  -999.0 \\
   l354pps &            & 18 00 50.0 & -23 20 33.0 &   6.55 &  -0.10 &  -999.0 \\
   l357pps &            & 17 40 57.3 & -31 10 48.0 & 357.56 &  -0.32 &  -999.0 \\
   l000pps &            & 17 46 37.1 & -29 10 21.5 & 359.91 &  -0.31 &  -999.0 \\
  l000pps2 &            & 17 47 10.3 & -28 46 11.8 &   0.32 &  -0.20 &  -999.0 \\
   l002pps &            & 17 50 36.3 & -27 06  2.8 &   2.14 &   0.01 &  -999.0 \\
   l006pps &            & 17 57 34.4 & -23 58 12.0 &   5.64 &   0.24 &   989.0 \\
   l009pps &            & 18 06 14.2 & -20 31 58.0 &   9.61 &   0.19 &  1187.0 \\
  l009pps2 &            & 18 06 15.5 & -20 32  8.0 &   9.61 &   0.19 &  1189.0 \\
   l012pps &            & 18 10 50.5 & -17 55 54.0 &  12.42 &   0.51 &  1205.0 \\
   l015pps &            & 18 14 35.3 & -16 45 45.0 &  13.87 &   0.28 &  1085.0 \\
  l015pps2 &            & 18 21 08.5 & -14 31 48.0 &  16.58 &  -0.05 &   748.0 \\
   l018pps &            & 18 25 42.0 & -13 10 13.0 &  18.30 &  -0.39 &  1544.0 \\
   l020pps &            & 18 28 09.9 & -11 28 47.0 &  20.08 &  -0.13 &  1650.0 \\
   l021pps &            & 18 30 34.0 &  -9 34 54.0 &  22.04 &   0.22 &   993.0 \\
   l023pps &            & 18 34 54.6 &  -8 49 18.0 &  23.20 &  -0.38 &  1681.0 \\
   l024pps &            & 18 36 05.1 &  -7 31 21.0 &  24.49 &  -0.04 &  1851.0 \\
   l025pps &            & 18 36 16.7 &  -6 43  8.7 &  25.23 &   0.29 &  -999.0 \\
   l026pps &            & 18 37 18.4 &  -6 38 38.4 &  25.41 &   0.10 &  -999.0 \\
  l026pps2 &            & 18 39 04.9 &  -6 24 23.6 &  25.82 &  -0.18 &  -999.0 \\
   l027pps &            & 18 41 51.1 &  -5 01 48.0 &  27.36 &  -0.17 &  3478.0 \\
   l028pps &            & 18 44 19.0 &  -4 40 52.0 &  27.95 &  -0.55 &  -999.0 \\
  l028pps2 &            & 18 42 51.8 &  -4 00  1.9 &  28.39 &   0.08 &  -999.0 \\
   l029pps &            & 18 42 15.6 &  -3 34 44.5 &  28.70 &   0.41 &  -999.0 \\
  l028pps2 &            & 18 42 50.8 &  -3 59 46.1 &  28.40 &   0.09 &  -999.0 \\
   l033pps &            & 18 52 25.1 &   0 14 53.0 &  33.26 &  -0.11 &  -999.0 \\
   l035pps &            & 18 53 39.1 &   1 50 32.0 &  34.82 &   0.35 &   756.0 \\
   l037pps &            & 18 59 10.0 &   4 12 10.0 &  37.55 &   0.20 &  1045.0 \\
   l038pps &            & 19 01 53.2 &   4 12 51.0 &  37.87 &  -0.40 &  1487.0 \\
   l040pps &            & 19 05 40.9 &   6 26  5.0 &  40.28 &  -0.22 &  1700.0 \\
   l042pps &            & 19 10 33.5 &   9 08  7.0 &  43.23 &  -0.05 &  1247.0 \\
   l044pps &            & 19 11 54.4 &   9 35 53.0 &  43.80 &  -0.13 &  1829.0 \\
  l044pps2 &            & 19 13 28.4 &  10 53 41.0 &  45.12 &   0.13 &  1999.0 \\
   l048pps &            & 19 23 10.8 &  14 26 31.0 &  49.37 &  -0.30 &  1903.0 \\
   l050pps &            & 19 23 11.1 &  14 26 34.0 &  49.37 &  -0.30 &  2088.0 \\
   l079pps &            & 20 29 25.3 &  40 11 28.0 &  78.89 &   0.71 &  1390.0 \\
   l079pps &            & 20 29 25.2 &  40 11 30.0 &  78.89 &   0.71 &  2014.0 \\
  l079pps2 &            & 20 31 13.3 &  40 03 33.0 &  78.99 &   0.35 &   790.0 \\
   l080pps &            & 20 34 42.8 &  39 44 45.0 &  79.13 &  -0.37 &  1104.0 \\
  l080pps2 &            & 20 40 05.3 &  41 31 48.0 &  81.17 &  -0.11 &   961.0 \\
   l080pps &            & 20 30 28.7 &  41 16  6.0 &  79.88 &   1.18 &  1024.0 \\
   l082pps &            & 20 40 26.8 &  41 56 54.0 &  81.54 &   0.10 &   805.0 \\
l110.5npps &            & 23 05 10.7 &  60 14 50.6 & 110.11 &   0.05 &  -999.0 \\
   l111pps &            & 23 15 31.5 &  61 07 39.7 & 111.62 &   0.38 &  -999.0 \\
    l134p1 &            & 02 29 02.8 &  61 33 31.2 & 134.28 &   0.86 &  -999.0 \\
  l134p1-2 &            & 02 27 06.9 &  61 52 20.8 & 133.95 &   1.07 &  -999.0 \\
  l134p1-3 &            & 02 27 02.4 &  61 52 32.9 & 133.94 &   1.07 &  -999.0 \\
    l135p1 &            & 02 34 45.1 &  61 46 23.0 & 134.83 &   1.31 &  -999.0 \\
   l136p15 &            & 02 50 08.4 &  61 59 56.7 & 136.38 &   2.27 &  -999.0 \\
   l137p15 &            & 02 29 03.2 &  60 43 26.9 & 134.59 &   0.08 &  -999.0 \\
    IC1396 &            & 21 40 11.5 &  58 16 11.7 &  99.93 &   4.21 &  -999.0 \\
  IC1396-2 &            & 21 36 36.8 &  57 30 50.0 &  99.07 &   3.97 &  -999.0 \\
  IC1396-3 &            & 21 35 38.4 &  57 26 40.5 &  98.93 &   4.00 &  -999.0 \\
   l189pps &            & 06 08 50.1 &  21 38 19.0 & 188.94 &   0.87 &  -999.0 \\
 l189pps-2 &            & 06 08 49.7 &  21 38  5.7 & 188.95 &   0.87 &  -999.0 \\
   l192pps &            & 06 12 51.5 &  18 00 39.3 & 192.58 &  -0.05 &  -999.0 \\
 l192pps-2 &            & 06 12 51.5 &  17 59 36.3 & 192.59 &  -0.05 &  -999.0 \\
 l192pps-3 &            & 06 12 50.1 &  18 00 35.8 & 192.58 &  -0.05 &  -999.0 \\
 l192pps-4 &            & 06 12 50.3 &  17 59 28.8 & 192.59 &  -0.06 &  -999.0 \\
 l192pps-5 &            & 06 07 43.9 &  20 39 29.6 & 189.67 &   0.17 &  -999.0 \\
}

\section{FITS Header Information}
\label{app:FITS_Header}

\begin{verbatim}

SIMPLE  =                    T / Written by IDL:  Sun Jan 18 01:21:05 2009      
BITPIX  =                  -32 / Number of bits per data pixel                  
NAXIS   =                    2 / Number of data axes                            
NAXIS1  =                 3134 /                                                
NAXIS2  =                  883 /                                                
DATE    = '2009-01-18'         / Creation UTC (CCCC-MM-DD) date of FITS header  
LONPOLE2=            180.00000 /lonpole                                         
LATPOLE2=            0.0000000 /latpole                                         
PPBEAM  =            21.276915 /pixels per beam                                 
CALIB_0 =          -0.00333379 / 0th coefficient for flux cal                   
CALIB_1 =             -2.92617 / 1st coefficient for flux cal                   
CALIB_2 =              6.97269 / 2nd coefficient for flux cal                   
ITERNUM =                    0 /Iteration number                                
N_PCA   =                    0 /number of PCA components subtracted             
COMMENT FITS (Flexible Image Transport System) format is defined in 'Astronomy  
COMMENT and Astrophysics', volume 376, page 359; bibcode 2001A&A...376..359H    
CTYPE1  = 'GLON-CAR'           / Coordinate Type                                
CTYPE2  = 'GLAT-CAR'           / Coordinate Type                                
EQUINOX =              2000.00 / Equinox of Ref. Coord.                         
CD1_1   =    -0.00199999986216 / Degrees / Pixel                                
CD2_1   =        0.00000000000 / Degrees / Pixel                                
CD1_2   =       -0.00000000000 / Degrees / Pixel                                
CD2_2   =     0.00199999986216 / Degrees / Pixel                                
CRPIX1  =              1270.52 / Reference Pixel in X                           
CRPIX2  =              366.367 / Reference Pixel in Y                           
CRVAL1  =        191.092003456 / Galactic longitude of reference pixel          
CRVAL2  =      0.0993870183708 / Galactic latitude of reference pixel           
PV2_1   =        0.00000000000 /Projection parameter 1                          
HISTORY PUTAST: Jan 18 01:10:42 2009 World Coordinate System parameters written 
END

\end{verbatim}

\include{tables}

\include{figures}

\bibliography{milkyway}

\end{document}